Finding Quality Issues in SKOS Vocabularies

### 2012-03-28 criteria.tex
minor changes

\todo{CM}{use definition environment and find out how to turn of numbering}
\paragraph{\textbf{Definition (SKOS Vocabulary).}}
Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV, CS \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C \subseteq IR\) with \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{concepts}, \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating concepts with one another, $LV \subseteq I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of untyped \textbf{plain literals}, and \(CS = I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I})\) being the set of \textbf{concept schemes}. Further, we let $V$ be the fully entailed RDFS interpretation of the underlying RDF graph, enriched by entailment of \texttt{owl:inverseOf} properties as well as instances of \texttt{owl:TransitiveProperty} and \texttt{owl:SymmetricProperty}.

@@

% Origin and rationale
The ISO/DIS 25964-1 standard \cite{ISO25964-1:2011} suggests that terms that share a common broader term but do not have an \todo{CM}{explain "overlapping"}overlapping meaning, should not be related associatively. This is advocated by Hedden~\cite{Hedden2010} and Aitchison et al.~\cite{Aitchison2000} who point out ``the risk that thesaurus compilers may overload the thesaurus with valueless relationships'', having a negative effect on precision.
% Function defintion
This issue can be checked by identifying the set of pairs of conceptual resources that share the same broader or narrower concept while also being associatively related by the property \texttt{skos:related}.

>>>>

\spnewtheorem*{mydef}{Definition}{\bfseries}{\itshape}
\begin{mydef}[SKOS Vocabulary]
Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV, CS \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C \subseteq IR\) with \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{concepts}, \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating concepts with one another, $LV \subseteq I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of untyped \textbf{plain literals}, and \(CS = I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I})\) being the set of \textbf{concept schemes}. Further, we let $V$ be the fully entailed RDFS interpretation of the underlying RDF graph, enriched by entailment of \texttt{owl:inverseOf} properties as well as instances of \texttt{owl:TransitiveProperty} and \texttt{owl:SymmetricProperty}. 
\end{mydef}

@@

% Origin and rationale
The ISO/DIS 25964-1 standard \cite{ISO25964-1:2011} suggests that terms that share a common broader term should not be related associatively if this relation is only justified by the fact that they are siblings. This is advocated by Hedden~\cite{Hedden2010} and Aitchison et al.~\cite{Aitchison2000} who point out ``the risk that thesaurus compilers may overload the thesaurus with valueless relationships'', having a negative effect on precision.
% Function defintion
This issue can be checked by identifying the set of pairs of conceptual resources that share the same broader or narrower concept while also being associatively related by the property \texttt{skos:related}.
